% This is samplepaper.tex, a sample chapter demonstrating the
% LLNCS macro package for Springer Computer Science proceedings;
% Version 2.20 of 2017/10/04
%
\documentclass[runningheads]{llncs}
%
\usepackage{graphicx}
% Used for displaying a sample figure. If possible, figure files should
% be included in EPS format.
\usepackage{subfig}
\usepackage{gensymb}
\usepackage{url}
\usepackage{amssymb}
\usepackage{multirow}

%
% If you use the hyperref package, please uncomment the following line
% to display URLs in blue roman font according to Springer's eBook style:
% \renewcommand\UrlFont{\color{blue}\rmfamily}

%for comments
%\newcommand{\FIXME}[1]{\textcolor{red}{FIXME: #1}}
\newcommand{\FIXME}[1]{{FIXME: #1}}

\begin{document}
%
\title{Computers Interacting with the Physical World:
A First-Year Course\thanks{The authors would like to acknowledge the
generous support of the Larsen family, which has been instrumental in the
development of this course.}
}

\titlerunning{Computers Interacting with the Physical World}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{Roger D. Chamberlain%\orcidID{0000-0002-7207-6106}
\and
Ron K. Cytron%\orcidID{1111-2222-3333-4444}
\and
Doug Shook%\orcidID{2222--3333-4444-5555}
\and
Bill Siever%\orcidID{0000-4444-5555-6666}
}
%
\authorrunning{R.D. Chamberlain et al.}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{Dept. of Computer Science and Engineering\\
Washington University in St. Louis, St. Louis, Missouri, USA\\
\email{\{roger,cytron,dshook,bsiever\}@wustl.edu}}
%

\maketitle              % typeset the header of the contribution
%
\begin{abstract}
Most introductory embedded systems offerings are upper-division courses.
At Washington Univ.~in St.~Louis,
embedded systems principles are introduced in a first-year course that
is required for both computer scientists and computer engineers.
% and currently has enrollments of 120-180 students each semester.
This paper describes the motivation for the course, its content,
the pedagogical techniques used, and lessons learned while developing
and administering the course.

\keywords{First-year course \and Cyber-physical systems.}
\end{abstract}
%
%
%

\input{intro}
\input{topics}
\input{delivery}
\input{weeks}
\input{lessons}
\input{conclude}

%\section*{Acknowledgements}
%The generous support of the Larsen family has been instrumental in the development of this course.
%The authors would like to acknowledge the generous support of the Larsen family, which has been instrumental in the development of this course.


%
% ---- Bibliography ----
%
% BibTeX users should specify bibliography style 'splncs04'.
% References will then be sorted and formatted in the correct style.
%
\bibliographystyle{splncs04}
\bibliography{paper}
%
\end{document}
